---
title: "STAT 341/641 Lab: Week Five"
author: "Firas Fakih"
date: "February 29 2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---
---

**STAT 341/641:**        Intro to EDA and Statistical Computing  
**Lab #5:**              Loops and the Jackknife  
**Teaching Assistant:**  "Erin Zhang"

* * *

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Directions:**  You will use the lab time to complete this assignment.

* * *

#**Task: Analyze the Gapminder Data**  

Install and load the gapminder data.
```{r}
#install.packages("gapminder")
library(gapminder)
library(dplyr)
library(ggplot2)
```


##1: Compute the average life expectancy and GDP per capita by country using the data from 1987 to 2007.  Plot these values in a scatterplot.  Color the points by the continent to which they belong. 

**Solution:**
```{r}
gapminder.new = subset(gapminder, year >= 1987 & year <= 2007)

gap.df = gapminder.new %>% group_by(country,continent) %>% summarise(meanLifeExp = mean(lifeExp),meanGDP = mean(gdpPercap))

ggplot(data = gap.df, aes(x = meanGDP,y = meanLifeExp)) + geom_point(aes(color = continent)) + ggtitle("Average GDP Per Capita vs Average Life Expectancy by Continent")
```
##2: Use the Mahalanobis distance method to detect outliers in the output from question one.  Which countries would you call outliers?  Why?

**Solution:**  
```{r}
vars <- gap.df[,3:4]
mu <- colMeans(vars)
covar <- cov(vars)
dists <- apply(vars, 1, mahalanobis, center = mu, cov = covar)
eta <- sqrt(qchisq(0.975, 2, 0))
mycols <- ifelse(dists < eta, "normal","outlier")
gap.df$Status = mycols
# Following countries are outliers
filter(gap.df, Status == "outlier") %>% select(country)

# They are Outliers because they are far away from the mean, so they have a low score on the mahalanobis distance scale
```

##3: Using the full dataset, implement the jackknife to identify which countries have a large influence on the $\beta_{\text{gdpPercap}}$ parameter in the regression equation
$$ lifeExp_i = \alpha + \beta_{\text{gdpPercap}}gdpPercap_i + \beta_{\text{pop}}pop_i + \epsilon_i.$$
To do this, write a for-loop.  In each iteration of the loop, drop one of the countries and compute the OLS coefficients.  Plot the value of $\beta_{\text{gdpPercap}}$ with the name of countries on the $x$-axis and the values of $\beta_{\text{gdpPercap}}$ on the $y$-axis.

**Solution:**  
```{r}
beta <- c()
countries <- gap.df$country
for (i in 1:142){
  data <- filter(gapminder,country != countries[i])
  beta[i] = lm(lifeExp ~ pop + gdpPercap, data)$coef[3]
}
plot(countries,beta)
identify(countries,beta,labels= countries)

# Kuwait has the largest influence on Beta
```

##5.4: Write a for-loop to sample $N=1,704$ points with replacement from the data.  Do this $R=250$ times.  For each iteration compute the mean of the population.  Visualize the 250 means with a boxplot.  Then compute the variance of the means.  Compare this to the standard estimate for the variance of the sample mean
$$\frac{\hat{\sigma}^2}{N}.$$

**Solution:**  
```{r}
N <- 1704
R <- 250
means <- numeric(R)
for (i in (1:R)){
  list <- sample_n(gapminder,size = N,replace = TRUE)
  means[i] = mean(list$pop)
}
boxplot(means)
var(means)
estimate <- (sd(means)^2)/N
estimate

# The standard estimate for the variance of the sample mean is smaller than the variance of the means.
```


* * *
